---
title: "Awards_Predictive_Analysis"
author: "Sekhar Mekala"
date: "Sunday, March 29, 2015"
output: 
 ioslides_presentation:
   widescreen: true
runtime: shiny
---



## Conditional probability model for Best Picture prediction

The main objectives of this data analysis are given below:

* Predict the probability of a picture getting the "Best Picture" award, given the picture is nominated in a specific award category

* Predict the probability of a picture getting the "Best Picture" award, given the picture won the award in a specific category

* Which academy award has the greatest influence on the Best Picture award.

The formula for conditional probability is given below:

<div class="blue">
**$$P(A|B) = P(B|A) . P(A)/P(B) = P(AB) / P(B)$$**
</div>


```{r, echo=FALSE,results='hide',cache.comments=FALSE,fig.align='center',warning=FALSE,message=FALSE}

#If you get error running this script, install 

#Reading the data from https://docs.google.com/spreadsheets/d/18YAYD2sVvErJziQGt9u2yitSZsjlDOJNuicTly0iaKw/edit?usp=sharing

#install.packages('gsheet')

library(gsheet)

library(dplyr)

library(tidyr)

awards_df <- gsheet2tbl("https://docs.google.com/spreadsheets/d/18YAYD2sVvErJziQGt9u2yitSZsjlDOJNuicTly0iaKw/edit?usp=sharing")
#head(awards_df)

awards_modified <- awards_df[,c(1,3,4,6)]

awards_modified <- awards_modified %>%
  group_by(movie_id,year,category_id) %>%
  summarise(won=max(won))

awards_re_modified <- spread((awards_modified),category_id,won)

award_categories <- unique(data.frame(category_id=awards_df$category_id,category_name=awards_df$category_name))
rownames(award_categories) <- NULL

#award_categories
names(awards_re_modified) <- c("movie_id", "year",  "c1",  "c2",  "c3",  "c4",  "c5", "c6",  "c7",  "c8",  "c9", "c10", "c11", "c12",
                               "c13", "c14", "c15","c16", "c17", "c18", "c19", "c20", "c21", "c22", "c23")

award_categories
awards_re_modified
```
We have the following data frames to use
* award_categories
* awards_re_modified

This R Markdown presentation is made interactive using Shiny. The viewers of the presentation can change the assumptions underlying what's presented and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Check box
```{r, echo=FALSE}
checkboxGroupInput("Nominated Categories", "Variable:",
                   c("Cylinders" = "cyl",
                     "Transmission" = "am",
                     "Gears" = "gear"))

vec <- as.vector(award_categories$category_name)
#names(vec) <- paste("c",1:23,sep="")

checkboxGroupInput("Nominated_Categories", "Nominated Categories:",
                   vec)

```
## Slide with Interactive Plot

```{r, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```


